---
title: "Julia for R"
author: "RCqls"
output: github_document
---
<!-- Rscript -e "rmarkdown::render('README.Rmd')" -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Julia for R

This is an attempt to embed the julia language in R. Actually, very basic julia types are converted to R objects. This package is also very experimental


## Install

1. Install Julia (all Operating System)

Install [Julia](https://julialang.org/downloads/). For Windows users don't forget to select `PATH` in the installer.

2. Windows user setup

* Install [RTools](https://cran.r-project.org/bin/windows/Rtools/) and a terminal with bash (for instance, [Git Bash](https://gitforwindows.org/))
* Ajout de Rscript dans le `PATH` (see for example [this page](https://www.hanss.info/sebastian/post/rtools-path/) for adding R path in the `PATH` environment variable)

3. Bash installation (all Operating Systems)

In a terminal (tested on macOS M1 with julia-1.9.2:) with `julia` and `Rscript` binaries supposed to be in the `PATH` environment variable,

```{bash, eval=FALSE}
/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/rcqls/jl4R/HEAD/inst/install.sh)"
```

## Test

Inside the R console:
## Examples

### getting started

### Goal: conversion of `julia` structures used in statitictic to `R`

* `DataFrame`

```{r}
require(jl4R)
jlusing(DataFrames)
jl(1)
interactive()
jl("(a=1,b=DataFrame(a=1:3,b=2:4))") -> nt_jl
nt_jl
toR(nt_jl)
R(nt_jl)
jl_typeof(nt_jl)
typeof(nt_jl)
```

* `CategoricalArray`

```{r}
jlusing(CategoricalArrays)
ca_jl <- jl('categorical(["titi","toto","titi"])')
ca_jl
R(ca_jl)
```